---
title: "Obtaining meteorology data from rnoaa package"
output:
  html_document:
    code_folding: hide
    code_download: true


---


```{r setup, incldue = FALSE, warning = FALSE, message = FALSE, echo = FALSE}
library(tidyverse)
setwd('../data/')
library(rnoaa)
korea = read.csv('south_korea.csv')
```

In this tutorial, we would like to show how we obtained the weather data through rnoaa package.
As a demonstration, we would like to show how we collected data for <tt>korea.csv</tt>. Let's take a look at the first several rows of the data.


```{r, results='asis'}
knitr::kable(head(korea))
```


Given latitude and longitude information, we would like to find weather stations, which have historical weather data.
To find the weather stations, first we need to get stations information using <tt>ghcnd_stations</tt> function from <tt>rnoaa</tt> package.
The data retrieved contains the basic information of each weather stations.

```{r, warning = FALSE, message = FALSE}
stations <- ghcnd_stations()
knitr::kable(head(stations))
```


There is another function, <tt>meteo_nearby_stations</tt>, which allows us to find nearby station given latitude and longitude information.
However, to use the function, it requires a dataframe with the certain column names; 'id','latitude', and 'longitude'.
Hence, we will first select the first three columns from korea and name it <tt>korea_latlong</tt> and change the column names.
Then we select the unique set of longitude and latitude to avoid searching for the same location multiple times.
```{r}
korea_latlong = korea[,1:3]
colnames(korea_latlong) = c('id','latitude','longitude')
korea_latlong = korea_latlong %>% unique()
knitr::kable(head(korea_latlong))
```


Now passing the <tt>korea_latlong</tt> into the <tt>meteo_nearby_stations</tt>, you can find a list of available stations for each of the latitude and longitude value.
Not all cities have a weather station in 20km radius. Among the cities with available weather station, we will show you Yeosu as an example. 
```{r}
result = meteo_nearby_stations(korea_latlong, station_data = stations, radius = 20, year_min = 1990) 
knitr::kable(result['South Korea/Yeosu'])
```

The result shows that there is one station 1.1km away from the latitude and longitude we have for the city of Yeosu.
Availability of the nearby weather station does not assure that the weather station store information we would like to explore.
In our case, we want to obtain 'tavg','tmin','tmax', and 'prcp'. Due to our imputation method, it is okay to contain two of the temperature variables.
Using for-loop, we will explore each of the weather station and check if they provide at least three of the four variables.
After the for-loop, it seems like the station near Yeosu has met our criteria.
```{r}
for(i in 1:length(result)){
  output = NULL
  if(dim(result[[i]])[1] > 1){
    station_ids = result[[i]]$id
    for (stationid in station_ids){
      list_df = ghcnd_search(stationid = stationid, var = c("tmax",'tmin','prcp','tavg'),
                             date_min = "1951-10-01", date_max = "2022-02-28")
      N = attributes(list_df)$names %>% length()
      N
      if(N>3){
        output = c(output, stationid)
      }
    }
    if(length(output) == 0){
      result[i] = NULL
    }
    else{
      result[[i]] = result[[i]][result[[i]]$id == output[1],]
    }
  }
}
knitr::kable(result['South Korea/Yeosu'])
```

Now back to korea_latlong table, we would like to narrow down to the list of cities where nearby weather station, which meets our criteria, exists.
To do so, we created 'exist' variable which indicates whether the weather station is available and selected only those rows that are TRUE for the variable.
We also add a variable 'stationid' which indicates the unique weather station id in the 'result' dataframe for each of the corresponding cities.
Finally, using the right merge, in the 'korea' dataframe, we have only the cities with available weather station and its weather station id.
```{r}
korea_latlong['exist'] = sapply(korea_latlong$id, function(x){dim(result[[as.character(x)]])[1] > 0}) %>% unlist()
korea_latlong = korea_latlong[korea_latlong$exist, ]
korea_latlong['stationid']  = sapply(korea_latlong$id, function(x){result[[as.character(x)]]$id   })
# we need to change the column names of the korea_latlong
colnames(korea_latlong)[1:3] = c('location', 'lat','long')
# merge
korea = left_join(x= korea, y = korea_latlong, by = c('location','lat','long')) %>% na.omit() %>% 
  dplyr::select(c('stationid','location','lat','long','alt','year','bloom_date','bloom_doy'))
knitr::kable(head(korea))
```

Now, going over each of the city, we will collect weather information from the corresponding weather stations defined by the 'stationid' column.
We will create a dataframe, 'output', which contain all the weather information between 1951 and early 2022.
Even though we set the begin date on October 1951, it returns either this or the earliest avaiable date.
```{r, warning = FALSE, message = FALSE}
output = NULL
for( stationid in unique(korea$stationid)){
  list_df = ghcnd_search(stationid = stationid, var = c("tmax",'tmin','prcp','tavg'), 
                         date_min = "1951-10-01", date_max = "2022-02-28")
  
  N = attributes(list_df)$names %>% length()
  df = list_df[[1]]
  df
  N
  if(N > 1){
  for(i in 2:N){
    df = left_join(x = df, y = list_df[[i]], by = c('id','date'))
  }}
  #df = df %>% dplyr::select(c('id','date','tmax','tmin','prcp'))
  
  df = df %>% dplyr::select(c('id','date',attributes(list_df)$names))
  colnames(df)[1] = 'stationid'
  output = plyr::rbind.fill(output, df)
}
knitr::kable(head(output))

```

Before merging this table with the 'korea' table, we need to mutate the variables.
December is going to be 0 and October and November of the previous year is going to be -2 and -1 respectively.
Also, for October, November, and December, their year value increase by 1.

```{r}
output = output %>% mutate(year = as.integer(format(date, "%Y")),
                           month = as.integer(strftime(date, '%m')) %% 12,
                           month = if_else(month %in% c(10, 11), month -12, month),
                           year = if_else(month <= 0, year + 1L, year))
knitr::kable(head(output))
```


Finally, using inner_join, we obtain the final table.
```{r}
final = inner_join(x = korea, y = output, by = c('stationid','year'))

knitr::kable(head(final))
```

