---
title: "Cherry blossom: Imputation of NA"
output: rmarkdown::github_document

---

<!--   rmarkdown::github_document
    code_folding: hide
    code_download: true -->



```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}

########### Data imputation

library(readr)
korea_complete<- read_csv("C:/Users/cheap/Downloads/korea_complete.csv")
kyoto_complete <- read_csv("C:/Users/cheap/Downloads/kyoto_complete.csv")
liestal_complete <- read_csv("C:/Users/cheap/Downloads/liestal_complete.csv")
washingtondc_complete <- read_csv("C:/Users/cheap/Downloads/washingtondc_complete.csv")
japan_complete <- read_csv("C:/Users/cheap/Downloads/japan_complete.csv")
us_complete <- read_csv("C:/Users/cheap/Downloads/us_complete.csv")
switzerland_complete <- read_csv("C:/Users/cheap/Downloads/switzerland_complete.csv")


```

In this page, I will demonstrate how we imputed NAs in our dataset. Let's see first few rows of <tt>korea_complete</tt> dataset. 

There are several <tt>NA</tt>s in <tt>tmin</tt> and <tt>prec</tt> data. We would like to replace these <tt>NA</tt>s with reasonable numbers. 

```{r, results='asis'}
knitr::kable(head(korea_complete))
```


We divided cases of missing data:

<b>1) When one of <tt>tmax</tt>, <tt>tmin</tt>, and <tt>tavg</tt> is missing. </b>

We replaced the missing values with regression prediction. For instance, to replace the following missing values, we ran linear regression with <tt>tmin</tt> on <tt>tmax</tt> and <tt>tavg</tt>. In the following case, we replaced the first <tt>NA</tt> of <tt>tmin</tt> with

$$tmin=-20.375867-0.505408 \times tmax+1.473154\times tavg$$. 

We did the same regression to replace <tt>tmax</tt> and <tt>tavg</tt>. We used the data of each country when we regress. 

```{r, results='asis'}
knitr::kable(korea_complete[c(7,8),])
```

```{r}
summary(lm(data=korea_complete, tmin~tmax+tavg))
```

<b>2) When two values are missing in one row.</b> Let's see the following case. 
In the following case, both <tt>tmax</tt> and <tt>tmin</tt> are missing. In this case, we used replaced the missing value with moving average method. In the case below, the <tt>tmax</tt> was replaced by $(236+190)/2=213$. We used moving average 4, and used exponential weights. 


```{r, results='asis'}
knitr::kable(korea_complete[c(30229:30231),])
```


<b>3) When <tt>prec</tt> is missing, we replaced it with 0. </b>

<tt>Temp\$prcp[is.na(Temp\$prcp)]<-0</tt>

```{r, eval=FALSE}
############## Imputation of NAs
filenames<-c("korea_complete","kyoto_complete","switzerland_complete", "washingtondc_complete", "japan_complete", "us_complete")

for(k in 1:length(filenames)){
  Temp<-get(filenames[k]) # define dataset here
  print(paste0(filenames[k], " Start!"))

  ## Check if the variable has different name.
  tmin.v<-Temp$tmin
  tmax.v<-Temp$tmax
  tavg.v<-Temp$tavg
  null.v<-c(is.null(tmin.v), is.null(tmax.v), is.null(tavg.v))
  if(sum(null.v)>0){
    print(paste0(filenames[k], " was not executed"))
  }else{
    # regress tmax, tmin, tavg on the others
    md.max<-lm(data=Temp, tmax~tmin+tavg)
    md.min<-lm(data=Temp, tmin~tmax+tavg)
    md.avg<-lm(data=Temp, tavg~tmax+tmin)
    
    for(i in 1:nrow(Temp)){
      df<-Temp[i,]
      tmin<-df$tmin
      tmax<-df$tmax
      tavg<-df$tavg
      NA.arr<-is.na(c(tmin, tmax, tavg))
      if(sum(NA.arr==c(1,0,0))==3){
        # replace missing tmin with prediction from the regression
        tmin<-predict(md.min,data.frame(tmax=tmax,tavg=tavg))
      }
      if(sum(NA.arr==c(0,1,0))==3){
         # replace missing tmin with prediction from the regression
        tmax<-predict(md.max,data.frame(tmin=tmin,tavg=tavg))
      }
      if(sum(NA.arr==c(0,0,1))==3){
         # replace missing tmin with prediction from the regression
        tavg<-predict(md.avg,data.frame(tmax=tmax,tmin=tmin))
      }
      Temp$tmin[i]<-tmin
      Temp$tmax[i]<-tmax
      Temp$tavg[i]<-tavg
      if(i%%10000==0){
        print(i)
      }
    }
    ## impute missing variables with MA(4), with exponential weight
    Temp$tavg<-imputeTS::na_ma(Temp$tavg)
    Temp$tmin<-imputeTS::na_ma(Temp$tmin)
    Temp$tmax<-imputeTS::na_ma(Temp$tmax)
    ## impute missing variable when prcp is missing. 
    Temp$prcp[is.na(Temp$prcp)]<-0
    ## write out the processed dataset
    write_csv(Temp, paste0(filenames[k], "_na_imputed.csv"))
    print(paste0(filenames[k], " Finished!"))
  }
}

```


<b>4) For Liestal, we could not attain <tt>tavg</tt> at all. </b>
Hence, we averaged <tt>tmin</tt> and <tt>tmax</tt> to approximate <tt>tavg</tt>. 

<tt>Temp\$tavg<-0.5\*Temp$tmax+0.5\*Temp\$tmin</tt>

```{r, eval=FALSE}
##################liestal
filenames<-"liestal_complete"
  Temp<-get(filenames) # define dataset here
  print(paste0(filenames[k], " Start!"))
  
  Temp$tavg<-0.5*Temp$tmax+0.5*Temp$tmin
  md.max<-lm(data=Temp, tmax~tmin+tavg)
  md.min<-lm(data=Temp, tmin~tmax+tavg)
  md.avg<-lm(data=Temp, tavg~tmax+tmin)
  
  #summary(md.max)
  
  for(i in 1:nrow(Temp)){
    df<-Temp[i,]
    tmin<-df$tmin
    tmax<-df$tmax
    tavg<-df$tavg
    NA.arr<-is.na(c(tmin, tmax, tavg))
    if(sum(NA.arr==c(1,0,0))==3){
      tmin<-predict(md.min,data.frame(tmax=tmax,tavg=tavg))
    }
    if(sum(NA.arr==c(0,1,0))==3){
      tmax<-predict(md.max,data.frame(tmin=tmin,tavg=tavg))
    }
    if(sum(NA.arr==c(0,0,1))==3){
      tavg<-predict(md.avg,data.frame(tmax=tmax,tmin=tmin))
    }
    Temp$tmin[i]<-tmin
    Temp$tmax[i]<-tmax
    Temp$tavg[i]<-tavg
    Temp$prcp[is.na(Temp$prcp)]<-0
    if(i%%10000==0){
      print(i)
    }
  }
  ## impute missing variables with MA(4), with exponential weight
  
  Temp$tavg<-imputeTS::na_ma(Temp$tavg)
  Temp$tmin<-imputeTS::na_ma(Temp$tmin)
  Temp$tmax<-imputeTS::na_ma(Temp$tmax)
  
  write_csv(Temp, paste0(filenames[k], "_na_imputed.csv"))
  print(paste0(filenames[k], " Finished!"))
  
```  




